---
title: "Digital Capacities Index - Quantitative Findings"
author: "Liam Magee, Delphine Bellarose, Anjali Sharma"
date: "31 March 2016"
output: html_document
---

<!-- output: word_document -->
<!-- output: pdf_document -->


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir="..")
knitr::opts_chunk$set(echo = FALSE)
setwd("..")

library(captioner)
library(pryr)
fig_nums <- captioner(prefix = "Figure")
fig_nums(name = "plot1", caption = "1st plot.")
fig_nums(name = "plot2", caption = "2nd plot.")
cf <- partial(fig_nums, display = "cite")

source("R/main.R", FALSE)
PRINTING <- FALSE

```

```{r}
# Set to false for non-HTML (e.g. word, pdf) outputs
PLOTLY <- FALSE
rmd_output <- tryCatch({rmarkdown::metadata$output},
                         error = function(e) {NULL})
if (rmd_output == 'html_document') {
  PLOTLY <- TRUE
}

print(PLOTLY)
```


## Digital Capacities Index - Test


### Sum of scores

The sum of the average scores to *interest* questions is: `r sum(colMeans(augmented.data[,vars.interest]))`.

The sum of the average scores to *competency* questions is: `r sum(colMeans(augmented.data[,vars.competencies]))`.

The sum of the average scores to *resilience* questions is: `r sum(colMeans(augmented.data[,vars.resilience]))`.

The sum of the average scores to *connectedness* questions is: `r sum(colMeans(augmented.data[,vars.connectedness]))`.

The sum of the average scores to all questions is: `r sum(colMeans(augmented.data[,vars.index]))`.


### Correlations


The correlation of *interests* to *competencies* is: `r cor(rowSums(augmented.data[,vars.interest]), rowSums(augmented.data[,vars.competencies]))`.

The correlation of *interests* to *resilience* is: `r cor(rowSums(augmented.data[,vars.interest]), rowSums(augmented.data[,vars.resilience]))`.

The correlation of *interests* to *connectedness* is: `r cor(rowSums(augmented.data[,vars.interest]), rowSums(augmented.data[,vars.connectedness]))`.

The correlation of *competencies* to *resilience* is: `r cor(rowSums(augmented.data[,vars.competencies]), rowSums(augmented.data[,vars.resilience]))`.

The correlation of *competencies* to *connectedness* is: `r cor(rowSums(augmented.data[,vars.competencies]), rowSums(augmented.data[,vars.connectedness]))`.

The correlation of *resilience* to *connectedness* is: `r cor(rowSums(augmented.data[,vars.resilience]), rowSums(augmented.data[,vars.connectedness]))`.

```{r echo=FALSE}
fn = local({
  i = 0
  function(x) {
    i <<- i + 1
    paste('Figure ', i, ': ', x, sep = '')
  }
})
```

```{r echo=FALSE}
plot(1:10)
```

**`r fig_nums("plot1")`**


```{r echo=FALSE}
plot(rnorm(10))
```

**`r fig_nums("plot2")`**


Referencing `r cf("plot2")`, then `r cf("plot1")`.