---
title: "Digital Capacities Index - Quantitative Findings"
author: "Liam Magee, Delphine Bellarose, Anjali Sharma"
date: "31 March 2016"
output: html_document
bibliography: DCI.bib
---

<!-- output: word_document -->
<!-- output: pdf_document -->


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir="..")
knitr::opts_chunk$set(echo = FALSE)
setwd("..")

source("R/main.R", FALSE)
PRINTING <- FALSE

```




## Digital Capacities Index - Quantitative Findings

The *Digital Capacities Index* is a pilot survey instrument developed by researchers at Western Sydney University and Google Australia. 
The survey was administered by [Pure Profile](https://www.pureprofile.com/us/) in Febuary 2016.


### Introduction

We included a total of `r length(vars.all)` items measuring (a) frequency of various online behaviour, (b) levels of agreement with statements about digital capacities, (c) perceived importance of online activities and (d) ease of use of digital technologies.

We further distinguished questions into the following key thematic areas, or what we have termed, following @james2014urban, 'critical issues'. These issues are:

 - Interests
 - Competencies
 - Resilience
 - Social Connectedness

These four issues were distilled from a list of nine issues that also included *Engagement*, *Inclusion*, *Policy Environment*, *Infrastructure* and *Consequences*. 

Against these four issues, we selected items and scales from existing sources in the literature where possible. In particular we drew from 'Kids Online'
[@livingstone2010risks], Helsper's [-@helsper2012corresponding] 'Corresponding fields model', a study by Humphry [-@humphry2014importance] of mobile use among homeless populations, and indicators compiled by the [Young and Well CRC](http://www.youngandwellcrc.org.au/). Other indicators were developed by the *Digital Capacities Index* team. 

A large number of candidate scales were distilled down to the list after two day-long workshops, and testing of the survey.



### Demographics


The survey included a total of `r format(sampleSize(), big.mark   = ",")` participants. We requested the survey provider provide a panel in terms of age groups, gender and geographic regions. As the panel provider recruited participants online, our pilot sample is expected to be skewed towards Australian citizens and families with comparatively high digital capacities. This caveat is signficant to the interpretation of our results below.

[Need something on family composition]


#### Age

Participant ages ranged from `r min(data$age)` to `r max(data$age)`, with a median value of `r median(data$age)`. 

*Graph 1* provides more detailed age demographics:

```{r ageFreq, echo=FALSE}

ageChart(data)

```


#### Gender

Participant gender is roughly evenly distributed. The survey included `r format(length(which(data$gender == "Female")), big.mark   = ",")` (`r format(100 * length(which(data$gender == "Female")) / length(data$gender), digits = 2)`%) women; `r format(length(which(data$gender == "Male")), big.mark   = ",")` (`r format(100 * length(which(data$gender == "Male")) / length(data$gender), digits = 2)`%) men; and `r format(length(which(data$gender == "Other")), big.mark   = ",")` (`r format(100 * length(which(data$gender == "Other")) / length(data$gender), digits = 2)`%) identifying as 'Other'.

Gender demographics are distributed as per *Graph 2*:

```{r genderFreq, echo=FALSE}

genderChart(data)

```



#### Combined Age and Gender


Combined age and gender demographics are distributed as per *Graph 3*:

```{r genderAndAgeFreq, echo=FALSE}

genderAndAgeChart(data)

```

These figures approximate to Australia's adult age distribution, as reported by the ABS in 2014 in *Graph 4* below, though with a considerably higher skew towards younger women and older men.


![Population Structure, Age and sex - Australia - 1994 and 2014](http://www.abs.gov.au/ausstats/abs@.nsf/0/1cd2b1952afc5e7aca257298000f2e76/Body/0.2A9C!OpenElement&FieldElemFormat=gif)


```{r ageFreqForEase, echo=FALSE}

generateSingleAgeFrequency(1, vars.ease, easeLabels)
```

[Other breakdowns - by urban/rural]



## Results by Critical Issue


### Interests


Q74:

```{r ageFreqForAggregate, echo=FALSE}

generateAgeFrequenciesForAggregate()
```



### Competencies

**TBD**


### Resilience

**TBD**


### Social Connectedness

**TBD**



## Overall results


<!--
```{r generateInterests, echo=FALSE}
generateInterests()
```

```{r generateCompetencies, echo=FALSE}
generateCompetencies()
```

```{r generateResilience, echo=FALSE}
generateResilience()
```

```{r generateConnectedness, echo=FALSE}
generateConnectedness()
```

```{r generateIndex, echo=FALSE}
generateIndex()
```

-->



## Appendex 1 - Digital Capacities Index Survey

[Include Full Survey Here]



## References


