---
title: "Digital Capacities Index - Quantitative Findings"
author: "Liam Magee, Delphine Bellarose, Anjali Sharma"
date: "31 March 2016"
output: html_document
bibliography: DCI.bib
---

<!-- output: word_document -->
<!-- output: pdf_document -->


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir="..")
knitr::opts_chunk$set(echo = FALSE)
setwd("..")

source("R/main.R", FALSE)
PRINTING <- FALSE

```




## Digital Capacities Index - Quantitative Findings

The *Digital Capacities Index* is a pilot survey instrument developed by researchers at Western Sydney University and Google Australia. 
The survey was administered by [Pure Profile](https://www.pureprofile.com/us/) in Febuary 2016.


### Demographics

The survey included a total of `r sampleSize()` participants. 

We included a total of `r length(vars.all)` items measuring (a) frequency of various online behaviour, (b) levels of agreement with statements about digital capacities, (c) perceived importance of online activities and (d) ease of use of digital technologies.

We further distinguished questions into the following key thematic areas, or what we have termed, following @james2014urban, 'critical issues'. These issues are:

 - Interests
 - Competencies
 - Resilience
 - Social Connectedness

These four issues were distilled from a list of nine issues that also included *Engagement*, *Inclusion*, *Policy Environment*, *Infrastructure* and *Consequences*. 

We also selected items and scales from existing sources in the literature where possible. In particular we drew from 'Kids Online'
[@livingstone2010risks], Helsper's [-@helsper2012corresponding] 'Corresponding fields model', a study by Humphry [-@humphry2014importance] of mobile use among homeless populations, and indicators compiled by the [Young and Well CRC](http://www.youngandwellcrc.org.au/). Other indicators were developed by the *Digital Capacities Index* team.



Age demographics are distributed as per Graph 1:

```{r ageFreq, echo=FALSE}

ageChart(augmented.data)

```


Gender demographics are distributed as per Graph 2:

```{r genderFreq, echo=FALSE}

genderChart(augmented.data)

```


Combined age and gender demographics are distributed as per Graph 3:

```{r genderAndAgeFreq, echo=FALSE}

genderAndAgeChart(augmented.data)

```

These figures approximates to Australia's adult age distribution, as reported by the ABS in 2014 in Graph 4 below, though with a considerably higher skew towards younger women and older men.


![Population Structure, Age and sex - Australia - 1994 and 2014](http://www.abs.gov.au/ausstats/abs@.nsf/0/1cd2b1952afc5e7aca257298000f2e76/Body/0.2A9C!OpenElement&FieldElemFormat=gif)


```{r ageFreqForEase, echo=FALSE}

generateSingleAgeFrequency(1, vars.ease, easeLabels)
```

## Aggregate

Q74:

```{r ageFreqForAggregate, echo=FALSE}

generateAgeFrequenciesForAggregate()
```

## Overall results

```{r generateInterests, echo=FALSE}
generateInterests()
```

```{r generateCompetencies, echo=FALSE}
generateCompetencies()
```

```{r generateResilience, echo=FALSE}
generateResilience()
```

```{r generateConnectedness, echo=FALSE}
generateConnectedness()
```

```{r generateIndex, echo=FALSE}
generateIndex()
```





## References


