---
title: "Digital Capacities Index - Quantitative Findings"
author: | 
  | Liam Magee^1^, Delphine Bellarose^1^, Anjali Sharma^1^, Louise Crabtree^1^, Philippa Collin^1^, Justine Humphry^1^, Paul James^1^, Emma Kearney^1^, Tanya Notley^1^, Amanda Third^1^, Samantha Yorke^2^
  | 1. Western Sydney University
  | 2. Google Australia
date: "4 April 2016"
bibliography: DCI.bib
output:
  html_document:
    toc: true
    toc_depth: 4
---

<!-- output: word_document -->
<!-- output: pdf_document -->


```{r setup, include=FALSE, echo=FALSE}
library(captioner)
library(pryr)

knitr::opts_knit$set(root.dir="..")
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
setwd("..")


source("R/main.R", FALSE)
PRINTING <- FALSE

# Set to false for non-HTML (e.g. word, pdf) outputs
PLOTLY <- FALSE
rmd_output <- tryCatch({rmarkdown::metadata$output},
                         error = function(e) {NULL})
if (rmd_output == 'html_document') {
  PLOTLY <- TRUE
}

# Set up figure numbering
fig_nums <- captioner(prefix = "Figure")
cf <- partial(fig_nums, display = "cite")
fig_nums("age.freq", "Age Frequency")
fig_nums("gender.freq", "Gender Frequency")
fig_nums("age.gender.freq", "Age & Gender Frequency")
fig_nums("age.gender.freq.abs", "Australia's Age & Gender Frequency (ABS 2014)")



```




## Digital Capacities Index - Quantitative Findings

The *Digital Capacities Index* is a pilot survey instrument developed by researchers at Western Sydney University and Google Australia. 
The survey was administered by [Pure Profile](https://www.pureprofile.com/us/) in Febuary 2016.


### Introduction

We included a total of `r length(vars.all)` items measuring (a) frequency of various online behaviour, (b) levels of agreement with statements about digital capacities, (c) perceived importance of online activities and (d) ease of use of digital technologies.

We further distinguished questions into the following key thematic areas, or what we have termed, following @james2014urban, 'critical issues'. These issues are:

 - **Competencies** (`r length(vars.competencies)` indicators). 
 - **Interests** (`r length(vars.interest)` indicators).
 - **Resilience** (`r length(vars.resilience)` indicators).
 - **Social Connectedness** (`r length(vars.connectedness)` indicators).

These four issues were distilled from a list of nine issues that also included *Engagement*, *Inclusion*, *Policy Environment*, *Infrastructure* and *Consequences*. 

Against these four issues, we selected items and scales from existing sources in the literature where possible. In particular we drew from 'Kids Online'
[@livingstone2010risks], Helsper's [-@helsper2012corresponding] 'Corresponding fields model', a study by Humphry [-@humphry2014importance] of mobile use among homeless populations, and indicators compiled by the [Young and Well CRC](http://www.youngandwellcrc.org.au/). Other indicators were developed by the *Digital Capacities Index* team. 

A large number of candidate scales were distilled down to the current list after two day-long workshops, and testing of the survey.



### Demographics


The survey included a total of `r format(sampleSize(), big.mark   = ",")` participants. We requested the survey provider provide a panel in terms of age groups, gender and geographic regions. As the panel provider recruited participants online, our pilot sample is expected to be skewed towards Australian citizens and families with comparatively high digital capacities. This caveat is signficant to the interpretation of our results below.


#### Age

Participant ages ranged from `r min(data$age)` to `r max(data$age)`, with a median value of `r median(data$age)`. 

*`r cf("age.freq")`* provides more detailed age demographics:

```{r ageFreq, echo=FALSE}

(gg <- chartWrap(ageChart(data)))

```

**`r fig_nums("age.freq")`**

These show participants' ages correspond approximately to Australia's adult demographic. `r top4AgesAsPercentage()` of participants were aged 35-54.



#### Gender

Participant gender is roughly evenly distributed. The survey included `r format(length(which(data$gender == "Female")), big.mark   = ",")` (`r format(100 * length(which(data$gender == "Female")) / length(data$gender), digits = 2)`%) women; `r format(length(which(data$gender == "Male")), big.mark   = ",")` (`r format(100 * length(which(data$gender == "Male")) / length(data$gender), digits = 2)`%) men; and `r format(length(which(data$gender == "Other")), big.mark   = ",")` (`r format(100 * length(which(data$gender == "Other")) / length(data$gender), digits = 2)`%) identifying as 'Other'.

Gender demographics are distributed, as show in `r cf("gender.freq")`:

```{r genderFreq, echo=FALSE}

(gg <- chartWrap(genderChart(data)))

```

**`r fig_nums("gender.freq")`**


#### Combined Age and Gender


Combined age and gender demographics are distributed as per *`r cf("age.gender.freq")`:

```{r genderAndAgeFreq, echo=FALSE}

(gg <- chartWrap(genderAndAgeChart(data)))

```

**`r fig_nums("age.gender.freq")`**

These figures approximate to Australia's adult age distribution, as reported by the ABS in 2014 in **cf("age.gender.freq.abs")** below, though with a considerably higher skew towards younger women and older men.


![Population Structure, Age and sex - Australia - 1994 and 2014](http://www.abs.gov.au/ausstats/abs@.nsf/0/1cd2b1952afc5e7aca257298000f2e76/Body/0.2A9C!OpenElement&FieldElemFormat=gif)

**fig_nums("age.gender.freq.abs")**



#### State and Location

Participants are distributed by state as follows:

```{r stateFreq, echo=FALSE}

(gg <- chartWrap(stateChart(data)))

```

The split of participants between urban and regional/rural is as follows:

```{r locationFreq, echo=FALSE}

(gg <- chartWrap(locationChart(data)))

```


## Results by Critical Issue



### Competencies

Our survey asked participants to respond to two questions about competencies:

 - Frequency of online activity
 - Perceived ease of conducting online activity


#### Frequency of online activity

*Frequency of online activity* measures frequency of 15 different activities, ranging from highly common activities such as sending email through to less common activities (in 2016), such as writing blogs.

##### By activity 

The graph below shows the relative frequencies of each activity. Using the Internet generally (for work, study, and for personal use), sending email and social networking are the most common activities. Streaming music, playing games with others, sharing media and writing blogs or diaries are comparatively uncommon activities.

```{r graphSubQuestion74, echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(
	vars.competencies.online.activities.74, 
	"Frequency", 
	frequencyLabels, 
	"online-activities-74")))

```


##### Aggregated online activity

```{r onlineActivities, echo=FALSE}

(gg <- chartWrap(histogram(vars.competencies.online.activities.74, 'Online Activities', 'gen/Q74')))

```

#### Perceived ease of conducting online activity

```{r competenciesEaseOfTasks, echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.competencies.431, "Ease", easeLabels, "competencies-ease-of-tasks-431")))

```


```{r ageFreqForEase, echo=FALSE}

generateSingleAgeFrequency(1, vars.ease, easeLabels)

```




### Interests


#### General Interests

```{r interestsGeneral, echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.interests.general.437, "Frequency", frequencyMonthLabels, "interests-general-437")))

```


#### Interest in seeking difference


```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.interests.difference.seeking.341, "Agreement", agreementLabels, "interests-difference-seeking-341")))

```


#### Interest in fitness and health improvement

```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.interests.fitness.352, "Frequency", frequencyMonthLabels, "interests-fitness-352")))

```


```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.interests.health.improvement.353, "Agreement", agreementLabels, "interests-health-improvement-353")))

```


#### Interest in keeping in touch

```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.interests.keeping.in.touch.430, "Importance", importanceLabels, "keeping-in-touch-430")))

```




### Resilience




#### Frequency of harmful events

```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.resilience.harm.events.434, "Frequency", frequencyMonthLabels, "resilience-harm-events-434")))

```


#### Responses to statements about online harms

```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.resilience.harms.agreement.435, "Agreement", agreementLabels, "resilience-harms-agreement-435")))

```


#### Willingness to engage with others

```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.resilience.engage.with.others.428, "Agreement", agreementLabels, "resilience-engage-with-others-428")))

```


### Social Connectedness

#### Maintaining connections

```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.connectedness.events.343, "Frequency", frequencyMonthLabels, "connectedness-events-343")))

```


```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.connectedness.maintenance.287, "Importance", importanceLabels, "maintaining-connections-287")))

```



#### Attitudes towards Technology


```{r echo=FALSE}

(gg <- chartWrap(graphSubQuestionFrequencies(vars.connectedness.tech.attitudes.429, "Agreement", agreementLabels, "connectedness-tech-attitudes-429")))

```





## Overall results - TESTING ONLY

### Index scores

```{r echo=FALSE}

(gg <- chartWrap(generateIndexChart()))

```


### Aggregated by age

#### Aggregated by age

```{r ageFreqForAggregate, echo=FALSE}

(gg <- chartWrap(generateAgeFrequenciesForAggregate()))

```

<!--
```{r generateInterests, echo=FALSE}
## generateInterests()
```

```{r generateCompetencies, echo=FALSE}
## generateCompetencies()
```

```{r generateResilience, echo=FALSE}
## generateResilience()
```

```{r generateConnectedness, echo=FALSE}
## generateConnectedness()
```

```{r generateIndex, echo=FALSE}
## generateIndex()
```

-->



## Appendex 1 - Digital Capacities Index Survey

[Include Full Survey Here]


## Appendex 2 - To Be Completed

 - Fix age axes - DONE
 - More detailed demographics
 - Fix References
 - Name all chunks
 - Name all figures
 



## References


