---
title: "Digital Capacities Index - Quantitative Findings"
author: "Liam Magee, Delphine Bellarose, Anjali Sharma"
date: "31 March 2016"
output: html_document
---

<!-- output: word_document -->
<!-- output: pdf_document -->


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir="..")
knitr::opts_chunk$set(echo = FALSE)
setwd("..")

source("R/main.R", FALSE)
PRINTING <- FALSE

```




## Digital Capacities Index - Quantitative Findings

The *Digital Capacities Index* is a pilot survey instrument developed by researchers at Western Sydney University and Google Australia. 
The survey was administered by Pure Profile in Febuary 2016.


### Demographics

The survey included a total of `r sampleSize()` participants.

Age demographics are distributed as per Graph 1:

```{r ageFreq, echo=FALSE}

ageChart(augmented.data)

```


Gender demographics are distributed as per Graph 2:

```{r genderFreq, echo=FALSE}

genderChart(augmented.data)

```


Combined age and gender demographics are distributed as per Graph 3:

```{r genderAndAgeFreq, echo=FALSE}

genderAndAgeChart(augmented.data)

```

These figures approximates to Australia's adult age distribution, as reported by the ABS in 2014 in Graph 4 below, though with a considerably higher skew towards younger women and older men.


![Population Structure, Age and sex - Australia - 1994 and 2014](http://www.abs.gov.au/ausstats/abs@.nsf/0/1cd2b1952afc5e7aca257298000f2e76/Body/0.2A9C!OpenElement&FieldElemFormat=gif)


```{r ageFreqForEase, echo=FALSE}

generateSingleAgeFrequency(1, vars.ease, easeLabels)
```

## Aggregate

Q74:

```{r ageFreqForAggregate, echo=FALSE}

generateAgeFrequenciesForAggregate()
```


